Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-02-06 paper.qmd
Found the solution for true model problem

true_model_df <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model.rds")
true_model_df_std <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model_std.rds")
wireframe_true_model <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model_connections.rds")
training_data_two_curvy_not_std <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_data_without_std.rds")

@@

data_two_curvy_not_std <- training_data_two_curvy_not_std |>
  mutate(type = "data")

true_model_two_curvy_std <- true_model_df_std |> 
  select(-ID) |> 
  mutate(type = "true model")

true_model_two_curvy <- true_model_df |> 
  select(-ID) |> 
  mutate(type = "true model")
```

@@

```{r}
#| label: two-curvy-true-proj
# Apply the scaling
df_model_data_two_curvy <- bind_rows(data_two_curvy_not_std, true_model_two_curvy)

@@

```{r}
#| label: model-prediction-umap-quollr

@@

<!-- langevitour with pD model for two-curvy clusters -->
```{r}
#| label: best-fit-umap1
#| fig-cap: "We start with a tSNE layout (a) of non-linear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (b) and then lifted to $4\\text{-}D$. A \\gD{} projection of the model fit from a tour through the $4\\text{-}D$ space is presented in c. Additionally, a \\gD{} projection of the fitted and true models (d) from the same tour demonstrates that the model accurately captures the non-linear structure of the data, although it does not fully represent the width of that structure. This highlights a key characteristic of tSNE, which compresses the data when applying tSNE. Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-parameter) choice. The default is used. Video of the langevitour animation is available at <>."
#| fig-width: 10
#| fig-height: 10

@@

```{r}
#| label: model-prediction-points-line

>>>>

true_model_df <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model.rds")
wireframe_true_model <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model_connections.rds")

@@

true_model_two_curvy <- true_model_df |> 
  mutate(type = "true model")
```

@@

```{r}
#| label: two-curvy-true-proj
# Apply the scaling
df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy)

@@

```{r}
#| label: model-prediction-umap-quollr
#| eval: false

@@

<!-- langevitour with pD model for two-curvy clusters -->
```{r}
#| label: best-fit-umap1
#| fig-cap: "We start with a tSNE layout (a) of non-linear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (b) and then lifted to $4\\text{-}D$. A \\gD{} projection of the model fit from a tour through the $4\\text{-}D$ space is presented in c. Additionally, a \\gD{} projection of the fitted and true models (d) from the same tour demonstrates that the model accurately captures the non-linear structure of the data, although it does not fully represent the width of that structure. This highlights a key characteristic of tSNE, which compresses the data when applying tSNE. Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)parameter choice. The default is used. Video of the langevitour animation is available at <>."
#| fig-width: 10
#| fig-height: 10

@@

```{r}
#| label: model-prediction-points-line
#| eval: false
